---
title: "Unsupervised Prediction"
output:
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## Key Ideas

- Sometimes you don't know the labels for prediction
- To build a predictor
    - Create clusters
    - Name clusters
    - Build predictor for clusters
- In a new dataset
    - Predict clusters
    
---

## `iris` Example Ignoring Species Labels

```{r}
data(iris)
library(ggplot2)
library(caret)
train.flags <- createDataPartition(y=iris$Species, p=0.7, list=F)
iris.train <- iris[train.flags,]
iris.test <- iris[-train.flags,]
dim(iris.train)
dim(iris.test)
```

---

## Cluster with k-means

```{r}
k.means1 <- kmeans(subset(iris.train, select=-c(Species)), centers=3)
iris.train$clusters <- as.factor(k.means1$cluster)
qplot(Petal.Width, Petal.Length, color=clusters, data=iris.train)
```

---

## Compare to real labels

```{r}
table(k.means1$cluster, iris.train$Species)
```

---

## Build Predictor

```{r}
fit <- train(clusters ~ ., data=subset(iris.train, select=-c(Species)), method="rpart")
table(predict(fit, iris.train), iris.train$Species)
```

---

## Apply on Test

```{r}
test.cluster.pred <- predict(fit, iris.test)
table(test.cluster.pred, iris.test$Species)
```

---

## Notes and Further Reading

- The `cl_predict` function in the `clue` package provides similar functionality
- Beware over-interpretation of clusters!
- This is one basic approach to [recommendation engines](https://en.wikipedia.org/wiki/Recommender_system)
- [Elements of Statistical Learning](http://www-stat.stanford.edu/~tibs/ElemStatLearn/)
- [Introduction to Statistical Learning](http://www-bcf.usc.edu/~gareth/ISL/)