---
title: "Bagging"
output:
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## Bootstrap Aggregating (Bagging)

**Basic Idea:**

1. Resample cases and recalculate predictions
2. Average or majority vote

**Notes**

- Similar bias
- Reduced variance
- More useful for nonlinear functions

---

## Ozone Data

```{r}
library(ElemStatLearn)
library(caret)
data(ozone, package="ElemStatLearn")
ozone <- ozone[order(ozone$ozone),]
head(ozone)
```

[http://en.wikipedia.org/wiki/Bootstrap_aggregating](http://en.wikipedia.org/wiki/Bootstrap_aggregating)

---

## Bagged Loess

```{r}
ll <- matrix(NA, nrow=10, ncol=155)
for(i in 1:10) {
    ss <- sample(1:dim(ozone)[1], replace=T)
    ozone0 <- ozone[ss,]
    ozone0 <- ozone0[order(ozone0$ozone),]
    loess0 <- loess(temperature ~ ozone, ozone0, span=0.2)
    ll[i,] <- predict(loess0, data.frame(ozone=1:155))
}
plot(ozone$ozone, ozone$temperature, pch=19, cex=0.5)
for(i in 1:10) {lines(1:155, ll[i,], col="grey", lwd=2)}
lines(1:155, apply(ll, 2, mean), col="red", lwd=2)
```

---

## Bagging in `caret`

- Some models perform bagging for you, in `train` function consider `method` options:
    - `bagEarth`
    - `treebag`
    - `bagFDA`
- Alternatively you can bag any model you choose using the `bag` function

---

## More Bagging in `caret`

```{r}
predictors <- data.frame(ozone=ozone$ozone)
temperature <- ozone$temperature
treebag <- bag(predictors, temperature, B=10,
               bagControl=bagControl(fit=ctreeBag$fit,
                                     predict=ctreeBag$pred,
                                     aggregate=ctreeBag$aggregate))
```

[http://www.inside-r.org/packages/cran/caret/docs/nbBag](http://www.inside-r.org/packages/cran/caret/docs/nbBag)

---

## Example of Custom Bagging (cont...)

```{r}
plot(ozone$ozone, temperature, col="lightgrey", pch=19)
points(ozone$ozone, predict(treebag$fits[[1]]$fit, predictors), pch=19,
       col="red")
points(ozone$ozone, predict(treebag, predictors), pch=19, col="blue")
```

---

## Parts of Bagging

```{r}
ctreeBag$fit
```

---

## Parts of Bagging

```{r}
ctreeBag$pred
```

---

## Parts of Bagging

```{r}
ctreeBag$aggregate
```

---

## Notes and Further Resources

**Notes**

- Bagging is most useful for nonlinear models
- Often used with trees -- an extension is random forests
- Several models use bagging in caret's `train` function

**Further Resources:**

- [Bagging](https://en.wikipedia.org/wiki/Bootstrap_aggregating)
- [Bagging and Boosting](https://stat.ethz.ch/education/semesters/FS_2008/CompStat/sk-ch8.pdf)
- [Elements of Statistical Learning](http://statweb.stanford.edu/~tibs/ElemStatLearn/)