---
title: "Basic Preprocessing"
output:
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## Why Preprocess?

```{r}
library(caret)
library(kernlab)
data(spam)
inTrain <- createDataPartition(y=spam$type, p=0.75, list=F)
training <- spam[inTrain,]
testing <- spam[-inTrain,]
hist(training$capitalAve, main="", xlab="ave. capital run length")
```

---

## Why Preprocess?

```{r}
mean(training$capitalAve)
sd(training$capitalAve)
```

---

## Standardizing

```{r}
trainCapAve <- training$capitalAve
trainCapAveS <- (trainCapAve - mean(trainCapAve)) / sd(trainCapAve)
mean(trainCapAveS)
sd(trainCapAveS)
```

---

## Standardizing - Test Set

```{r}
testCapAve <- testing$capitalAve
testCapAveS <- (testCapAve - mean(trainCapAve)) / sd(trainCapAve)
mean(testCapAveS)
sd(testCapAveS)
```

---

## Standardizing - `preProcess` Function

```{r}
preObj <- preProcess(training[,-58], method=c("center","scale"))
trainCapAveS <- predict(preObj, training[,-58])$capitalAve
mean(trainCapAveS)
sd(trainCapAveS)
```

---

## Standardizing - `preProcess` Function

```{r}
testCapAveS <- predict(preObj, testing[,-58])$capitalAve
mean(testCapAveS)
sd(testCapAveS)
```

---

## Standardizing - `preProcess` Argument

```{r}
set.seed(32343)
modelFit <- train(type ~ ., training, preProcess=c("center","scale"),
                  method="glm")
modelFit
```

---

## Standardizing - Box-Cox Transforms

```{r}
preObj <- preProcess(training[,-58], method=c("BoxCox"))
trainCapAveS <- predict(preObj, training[,-58])$capitalAve
par(mfrow=c(1,2))
hist(trainCapAveS)
qqnorm(trainCapAveS)
```

---

## Standardizing - Imputing Data

```{r}
set.seed(13343)

# Make some values NA
training$capAve <- training$capitalAve
selectNA <- rbinom(dim(training)[1], size=1, prob=0.05)==1
training$capAve[selectNA] <- NA

## Impute and standardize
preObj <- preProcess(training[,-58], method="knnImpute")
capAve <- predict(preObj, training[,-58])$capAve

## Standardize true values
capAveTruth <- training$capitalAve
capAveTruth <- (capAveTruth - mean(capAveTruth)) / sd(capAveTruth)
```

---

## Standardizing - Imputing Data

```{r}
quantile(capAve - capAveTruth)
quantile((capAve - capAveTruth)[selectNA])
quantile((capAve - capAveTruth)[!selectNA])
```

---

## Notes and Further Reading

- Training and test must be processed in the same way
- Test transformations will likely be imperfect
    - Especially if the test / training sets were collected at different times
- Careful when trainsforming factor variables!
- [Preprocessing with caret](https://topepo.github.io/caret/preprocess.html)