---
title: "Training Options"
output:
    html_document:
        keep_md: true
---

```{r setup, cache=FALSE, echo=FALSE, message=F, warning=F, tidy=FALSE}
require(knitr)
options(width=100)
opts_chunk$set(message=F, error=F, warning=F, comment=NA, fig.align='center', dpi=100, tidy=F, cache.path='.cache/', fig.path='fig/')

options(xtable.type='html')
knit_hooks$set(inline=function(x) {
    if(is.numeric(x)) {
        round(x, getOptions('digits'))
    } else {
        paste(as.character(x), collapse=', ')
    }
})
knit_hooks$set(plot=knitr:::hook_plot_html)
```

## SPAM Example

```{r}
library(caret)
library(kernlab)
data(spam)
inTrain <- createDataPartition(y=spam$type, p=0.75, list=F)
training <- spam[inTrain,]
testing <- spam[-inTrain,]
modelFit <- train(type ~ ., training, method="glm")
```

---

## Train Options

```{r}
args(train.default)
```

---

## Metric Options

**Continuous outcomes:**

- _RMSE_ -- Root mean squared error
- _RSquared_ -- $R^2$ from regression models

**Categorical outcomes:**

- _Accuracy_ -- Fraction correct
- _Kappa_ -- A measure of [concordance](https://en.wikipedia.org/wiki/Cohen%27s_kappa)

---

## `trainControl`

```{r}
args(trainControl)
```

---

## `trainControl` resampling

- _method_
    - _boot_ -- bootstrapping
    - _boot632_ -- bootstrapping with adjustment
    - _cv_ -- cross-validation
    - _repeatedcv_ -- repeated cross-validation
    - _LOOCV_ -- leave-one-out cross-validation
- _number_
    - For boot/cross-validation
    - Number of subsamples to take
- _repeats_
    - Number of times to repeat subsampling
    - If big this can _slow things down_

---

## Setting the Seed

- It is often usedful to set an overall seed
- You can also set a seed for each resample
- Seeding each resample is useful for parallel fits

---

## Seed Example

```{r}
set.seed(1235)
modelFit2 <- train(type ~ ., training, method="glm")
modelFit2
```

---

## Seed Example

```{r}
set.seed(1235)
modelFit3 <- train(type ~ ., training, method="glm")
modelFit3
```

---

## Further Resources

- [Caret tutorial](http://www.edii.uclm.es/~useR-2013/Tutorials/kuhn/user_caret_2up.pdf)
- [Model training and tuning](https://topepo.github.io/caret/training.html)